\chapter{Formal Concept Analysis}
\label{chapter:formal-concept-analysis}

\FCA (henceforth initialised as FCA) provides a mathematical framework for reasoning about concepts and corresponding conceptual structures. It is an example of applied lattice theory, and benefits from rich, rigorous mathematical theory, as well as an easily graspable intuition. 

Prior to any direct discussion on FCA, it is important to provide some contextual justification for the philosophical ideas of concepts that were adopted into the foundation of FCA. 

\FCA (henceforth initialised as FCA) provides a simple, yet mathematically rigorous, framework for identifying and reasoning about ``concepts'' and their corresponding hierarchies in data \cite{ganter1999formal,ganter2016conceptual}. 

The central view of concepts as a dual between \textit{extension}---what one refers to as instances of a concept---and \textit{intension}---what meaning is ascribed to a concept---is supported by a rich philosophical backing. \index{formal concept analysis} \index{extension} \index{intension}

\section{Basic Notions}
\label{section:basic-notions}
The universe of discourse in FCA is made-up of sets of \textit{objects} and \textit{attributes}. Objects are extensional, they are the things pointed to as instances of some more general concept. In turn, attributes construct the intensional component of a concept.

We collect these extensional and intensional building blocks in a structure called a \textit{formal context}, which includes a binary relation that allows traversal from extension to intension, and vice verse.

\begin{definition}
  \label{definition:formal-context}
  A \textit{formal context} $\Fcontext$ is a triple comprised of a set of objects $G$, a set of attributes $M$, and a binary relation $I \subseteq G \times M$ referred to as an `incidence' relation. For an object-attribute pair $(g,m) \in I$ we might say that \say{object $g$ \textit{has} the attribute $m$}.
\end{definition}

A formal context \index{formal context} in some sense describes an open-world interpretation, and so $(g,m) \not \in I$ is not usually interpreted as saying that \say{object $g$ has the negation of the attribute $m$}.

\begin{example}
  \label{example:first-example-formal-context}
  Finite formal contexts of a reasonable size can be described entirely by a tabular representation. Each object corresponds to a row, and each attribute to a column.
  \begin{figure}[H]
    \centering
    \scriptsize
    \begin{tabular}{|c||c|c|c|c|c|}
      \hline
      & \rotatebox{0}{\texttt{closure}}
      & \rotatebox{0}{\texttt{associativity}}
      & \rotatebox{0}{\texttt{identity}}
      & \rotatebox{0}{\texttt{divisibility}}
      & \rotatebox{0}{\texttt{commutativity}} \\
      \hline\hline
      \texttt{magma}             & $\times$ &          &          &          &          \\ \hline
      \texttt{semigroup}         & $\times$ & $\times$ &          &          &          \\ \hline
      \texttt{monoid}            & $\times$ & $\times$ & $\times$ &          &          \\ \hline
      \texttt{group}             & $\times$ & $\times$ & $\times$ & $\times$ &          \\ \hline
      \texttt{abelian group}     & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ \hline
      \texttt{loop}              & $\times$ &          & $\times$ & $\times$ &          \\ \hline
      \texttt{quasigroup}        & $\times$ &          &          & $\times$ &          \\ \hline
      \texttt{groupoid}          &          & $\times$ & $\times$ & $\times$ &          \\ \hline
      \texttt{category}          &          & $\times$ & $\times$ &          &          \\ \hline
      \texttt{semicategory}      &          & $\times$ &          &          &          \\ \hline
      \specialrule{1.25pt}{0pt}{0pt}
    \end{tabular}
    \caption{A formal context showing necessary properties of group-like structures.}
    \label{figure:formal-context-group-structures}
  \end{figure}
\end{example}

The tabular representation of a formal context allows for easy identification of the set of attributes that a given object satisfies: one need only scan across the respective row in the table and note where an `$\times$' symbol appears. This set of attributes is called the \textit{object intent}. The dual notion of an \textit{attribute extent} can similarly be found by scanning down the column of a given attribute. The utility of this visual metaphor diminishes when, as is often the case, we consider larger sets of objects or attributes. Instead, we opt for a more formal approach to determining the intents and extents for (sets of) objects and attributes, respectively. \index{derivation operators! object intent} \index{derivation operators! attribute extent}

\begin{definition}
  \label{definition:derivation-operators}
  Given a formal context $\fcontext$, the \textit{derivation operators} are two order-reversing maps $(\cdot)^\uparrow : 2^G \to 2^M$ and $(\cdot)^\downarrow : 2^M \to 2^G$ where the order is given by subset inclusion. Then, for any subsets $A \subseteq G$ and $B \subseteq M$,
  \begin{align*}
    A^\uparrow \coloneqq \{m \in M \mid \forall g \in A, \; \langle g,m \rangle \in I\} \\
    B^\downarrow \coloneqq \{g \in G \mid \forall m \in B, \; \langle g,m \rangle \in I\}
  \end{align*}
\end{definition}

The derivation operators provide a clear way of describing, for a given set $A\subseteq G$ of objects, the set of attributes which every object in $A$ satisfies, denoted $A^\uparrow$. As an illustration, given the set of objects $\{\texttt{semigroup, monoid}\}$ from \Cref{figure:formal-context-group-structures}, its derivation would be $\{\texttt{closure, associativity}\}$. It is quite easy to spot that this is just the intersection of the object intents of \texttt{semigroup} and \texttt{monoid}.

Of course, these two functions can be composed; and so, $A^{\uparrow \downarrow}$ would yield the set of objects 

$\{\texttt{semigroup, monoid, group, abelian group}\}$, which can rather cumbersomely be described as ``the set of all  objects that satisfy all the attributes satisfied by \texttt{semigroup} and \texttt{monoid}''.

% If we were to then take the derivation of this set of attributes, $A^{\uparrow \downarrow}$, we get the set of objects $\{\texttt{semigroup, monoid, group, abelian group}\}$: the set of all group-likes that have a closed and associative binary operation!

In fact, this double-application of derivation operators satisfies very specific properties,
\begin{align}
  & \text{(monotonicity)} & A \subseteq A_1 \textit{ implies } A^{\uparrow \downarrow} \subseteq A_1^{\uparrow \downarrow} \\
  & \text{(extensivity)}  & A \subseteq A^{\uparrow \downarrow} \\
  & \text{(idempotency)}  & A^{\uparrow \downarrow} = (A^{\uparrow \downarrow})^{\uparrow \downarrow}
\end{align}
for all $A, A_1 \subseteq G$. Thus, $(\cdot)^{\uparrow \downarrow}$ describes a closure operator on $2^G$. The dual notion holds for attributes, and $(\cdot)^{\downarrow \uparrow}$ describes a closure operator on $2^M$.

\begin{proposition}
  \label{proposition:properties-about-derivation-operators}
  Let $\fcontext$ be a formal context with subsets $A_0, A_1, A_2 \subseteq G$ and $B_0, B_1, B_2 \subseteq M$ of attributes. Then,
  \begin{center}
    \begin{minipage}[t]{0.48\textwidth}
      \begin{enumerate}
        \item $A_0 \subseteq A_1 \Rightarrow A_1^\uparrow \subseteq A_0^\uparrow$
        \item $A_0 \subseteq A_0^{\uparrow \downarrow}$
        \item $A_0^\uparrow = A_0^{\uparrow \downarrow \uparrow}$
      \end{enumerate}
    \end{minipage}%
    \hfill
    \begin{minipage}[t]{0.48\textwidth}
      \begin{enumerate}
        \item $B_0 \subseteq B_1 \Rightarrow B_1^\downarrow \subseteq B_0^\downarrow$
        \item $B_0 \subseteq B_0^{\downarrow \uparrow}$
        \item $B_0^\downarrow = B_0^{\downarrow \uparrow \downarrow}$
      \end{enumerate}
    \end{minipage}
  \end{center}
\end{proposition}

\begin{definition}
  \label{definition:formal-concept}
  A \textit{formal concept} of a formal context $\fcontext$ is a pair $(A,B)$ of subsets $A \subseteq G$ and $B \subseteq M$ that satisfies $A^\uparrow = B$ and $B^\downarrow = A$. Then, $A$ is the concept \textit{extent} and $B$ is the \textit{intent}. We write $\B$ to denote the set of all concepts of $\fcontext$.
\end{definition}
\index{formal concept}

Now explain why derivation of set of objects is always a concept intent by proposition 1. And the dual. and Galois connections
\index{Galois connections}

\begin{figure}[H]
  \centering
  \begin{tikzpicture}[every node/.style={circle, fill=black, inner sep=1pt}]
    \node (a1) at (0,3) [label=above:{\scriptsize $\top$}] {w};
    % \node[draw=none, fill=none] at (0,3.5) {\tiny Top};

    \node (b1) at (-3,1) [label=above:{\scriptsize associativity}] {y};
    \draw (a1) -- (b1);

    \node (b2) at (-1,1) [label=above:{\scriptsize identity}] {z};
    \draw (a1) -- (b2);

    \node (b3) at (1,1) [label=above:{\scriptsize closure},label=below:{\scriptsize magma}] {x};
    \draw (a1) -- (b3);

    \node (b4) at (3,1) [label=above:{\scriptsize divisibility}] {x};
    \draw (a1) -- (b4);

    \node (c1) at (-4,-1) [label=below:{\scriptsize category}] {y};
    \draw (c1) -- (b1);
    \draw (c1) -- (b2);

    \node (c2) at (-2,-1) [label=below:{\scriptsize semigroup}] {y};
    \draw (c2) -- (b1);
    \draw (c2) -- (b3);

    \node (c3) at (0,-1) {y};
    \draw (c3) -- (b2);
    \draw (c3) -- (b3);

    \node (c4) at (2,-1) {y};
    \draw (c4) -- (b2);
    \draw (c4) -- (b4);

    \node (c5) at (4,-1) [label=below:{\scriptsize quasigroup}] {y};
    \draw (c5) -- (b3);
    \draw (c5) -- (b4);

    \node (d1) at (-2,-3) [label=below:{\scriptsize monoid}] {y};
    \draw (d1) -- (c1);
    \draw (d1) -- (c2);
    \draw (d1) -- (c3);

    \node (d2) at (0,-3) [label=below:{\scriptsize groupoid}] {y};
    \draw (d2) -- (c1);
    \draw (d2) -- (c3);

    \node (d3) at (2,-3) [label=below:{\scriptsize loop}] {y};
    \draw (d3) -- (c3);
    \draw (d3) -- (c4);
    \draw (d3) -- (c5);

    \node (e1) at (0,-5) [label=below:{\scriptsize group}] {y};
    \draw (e1) -- (d1);
    \draw (e1) -- (d2);
    \draw (e1) -- (d3);

    \node (f1) at (0,-7)[label=above:{\scriptsize commutativity},label=below:{\scriptsize abelian group}] {y};
    \draw (f1) -- (e1);
  \end{tikzpicture}
  \caption{The concept lattice associated with the formal context in \Cref{figure:formal-context-group-structures}}
\end{figure}

